Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-13 thesis/chapters/ch_introduction.tex
rearranged sections a bit
updating introduction and references. Hydrodynamics section, more in opening intro (observational campaigns), and some in radiation section

@@

Galaxies are amalgamations of gas and stars embedded in dark matter halos, formed over cosmic time as the products of the hierarchical assembly that follows from the collapse of primordial density fluctuations that arose after the Big Bang. The gravitational pull of dark matter controls the growth of structure in the Universe which, as is predicted from $\Lambda$CDM cosmology, TEXT HERE. While, broadly speaking, the evolution of baryons is dominated by the gravitational pull of dark matter, the beautiful simplicity of the evolution of dark matter and $\Lambda$CDM is muddied by their existence. The complexities of hydrodynamics, magnetohydrodynamics, thermodynamics, chemistry, radiative processes, and nucleosynthesis -- or, ``astrophysics'' for short -- drives these deviations and gives rise to the Universe that we observe and interact with and -- not least of all -- us. In spite of a concerted effort spanning nearly a century, we are far from a self-consistent theory of galaxy evolution \citep[see ][ for a recent review of outstanding problems in galactic evolution]{NaabOstriker2017}. It is understanding the rich set of physics that formed our own Galaxy and the countless galaxies scattered throughout the Universe that motivates this work.

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, such as Li, all of the elements in the Universe are produced in nuclear reactions in the cores of stars, massive star stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like nuetron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{}, and detailed stellar abundances in galaxies with spectroscopically resolved stellar populations \citep[e.g.][]{}.

Developing a complete model for galactic chemical evolution is a daunting task. One is faced with the already non-trivial prospect nailing down a complete model of stellar structure, stellar evolution, and the rates of individual nuclear reactions. One must then perscribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas into the disk of the galaxy, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, how, and with what masses stars form in the first place from this gas, which requires understanding the hydrodynamic and magnetohydrodynamic properties of the multi-scale, multi-phase turbulent interstellar medium (ISM). This in turn requires one to a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and supernovae -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Once this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda CDM$ model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift.....

 It should be obvious now to the reader why there remains so much to be learned about galactic chemical evolution. Yet it makes clear that this field offers an incredibly exciting test of an incredible range of physics.

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns

CE models

@@

@@

Stars, particularly large-scale cosmological simulations, are commonly represented as simple stellar populations (SSPs), or particles with IMF-averaged properties of stellar feedback and metal enrichment. As demonstrated in \cite{Revaz2016}, this approach is reasonable so long as the mass of the star particle is large enough to fully sample the IMF ($> 10^{4}$~M$_{\odot}$). In high resolution simulations, particularly simulations of small, low-mass systems like dwarf galaxies, this is no longer the case. This leads to inconsistencies in both the effective metal yield of the modelled stars, and (often) an artificial reduction of the effects of stellar feedback. Recent works have developed new models to address this issue, discretizing (rather than averaging over) individual feedback events \citep[e.g.][]{Stinson2010,Hopkins2014,Hopkins2018,Rosdahl2018} or accounting for stochastic sampling of the IMF \citep{Hu2016,Hu2017,Applebaum2018,Su2018}. However, each of these methods utilizes a sub-grid recipe in some fashion to improve upon the SSPs. On the other extreme are highly resolved simulations which model star formation with ``sink particles'' \citep[see for example ][]{Krumholz2004,Federrath2010,GongOstriker2013,BleulerTeyssier2014,Sormani2017} that account for the gradual accretion of gas in the growth of a single star (or group of stars) and protostellar feedback. These recipes are often computationally expensive, however, restricting their use over long timescales (100's to 1000's of Myr or more) in galaxy-scale simulations.

@@

\subsubsection{Stellar Abundances}

\subsubsection{Sources of Enrichment}

\subsection{Modelling Chemical Evolution}

\subsubsection{Hydrodynamics Simulations}

\subsubsection{Semi-analytic Models / One-zone models}

@@

%
%
%

>>>>

\newcommand{\code}{\textsc}

@@

Galaxies are amalgamations of gas and stars embedded in dark matter halos, formed over cosmic time as the products of the hierarchical assembly that follows from the collapse of primordial density fluctuations that arose after the Big Bang. The gravitational pull of dark matter controls the growth of structure in the Universe which, as is predicted from $\Lambda$CDM cosmology, TEXT HERE. While, broadly speaking, the evolution of baryons is dominated by the gravitational pull of dark matter, the beautiful simplicity of the evolution of dark matter and $\Lambda$CDM is muddied by their existence. The complexities of hydrodynamics, magnetohydrodynamics, thermodynamics, chemistry, radiative processes, and nucleosynthesis -- or, ``astrophysics" for short -- drives these deviations and gives rise to the Universe that we observe and interact with and -- not least of all -- us. In spite of a concerted effort spanning nearly a century, we are far from a self-consistent theory of galaxy evolution \citep[see ][ for a recent review of outstanding problems in galactic evolution]{NaabOstriker2017}. It is understanding the rich set of physics that formed our own Galaxy and the countless galaxies scattered throughout the Universe that motivates this work.

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions in the cores of stars, massive star stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like nuetron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{}, and detailed stellar abundances in galaxies with spectroscopically resolved stellar populations \citep[e.g.][]{}.

%Developing a complete model for galactic chemical evolution is a daunting task.
Here is a recipe for constructing a complete model of galactic chemical evolution as it is understood today. One begins first with the simple step of nailing down a complete model of stellar structure, stellar evolution, and learning from nuclear physicists the reaction rates of every nuclear reaction and lifetimes of individual isotopes. One must then perscribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas into the disk of the galaxy, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, how, and with what masses stars form in the first place from this gas, which requires understanding the hydrodynamic 
%and magnetohydrodynamic 
properties of the multi-scale, multi-phase turbulent interstellar medium (ISM). This in turn requires one to toss in a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and supernovae -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Once this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda CDM$ model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift..... Wait... I forgot about magnetic fields and cosmic rays...

It should be obvious now to the reader why there remains so much to be learned about galactic chemical evolution. Developing a complete model for galactic chemical evolution is a daunting task. Yet it is clear that this field offers an incredibly exciting test of an incredible range of physics.

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundnaces and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeaology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain. The excitement at the availability of this data does not even mention

One of the most common ways to model and understand the stellar populations of galaxies in high-dimensional chemical space is through simplified one-zone (or many-zone) models. This treatment extends back to XXX, where galaxies were treated as closed-box systems of gas and stars with simple approximations such as the instantaneous recycling of metal yields from massive stars and homogenous mixing. While simple, these models have been used extensively to explain the mean evolution of stars in chemical space in individual galaxies.

@@

\section{Observations of Local Group Dwarf Galaxies}

@@

Stars, particularly large-scale cosmological simulations, are commonly represented as simple stellar populations (SSPs), or particles with IMF-averaged properties of stellar feedback and metal enrichment. As demonstrated in \cite{Revaz2016}, this approach is reasonable so long as the mass of the star particle is large enough to fully sample the IMF ($> 10^{4}$~M$_{\odot}$). In high resolution simulations, particularly simulations of small, low-mass systems like dwarf galaxies, this is no longer the case. This leads to inconsistencies in both the effective metal yield of the modelled stars, and (often) an artificial reduction of the effects of stellar feedback. Recent works have developed new models to address this issue, discretizing (rather than averaging over) individual feedback events \citep[e.g.][]{MUGS2010,FIRE,Hopkins2018,Rosdahl2018} or accounting for stochastic sampling of the IMF \citep{Hu2016,Hu2017,Applebaum2018,Su2018}. However, each of these methods utilizes a sub-grid recipe in some fashion to improve upon the SSPs. On the other extreme are highly resolved simulations which model star formation with ``sink particles" \citep[see for example ][]{Krumholz2004,Federrath2010,GongOstriker2013,BleulerTeyssier2014,Sormani2017} that account for the gradual accretion of gas in the growth of a single star (or group of stars) and protostellar feedback. These recipes are often computationally expensive, however, restricting their use over long timescales (100's to 1000's of Myr or more) in galaxy-scale simulations.

@@

We only briefly discuss this method of feedback here, as it is discussed in much more detail in both Chapters~\ref{ch:chapter_1} and \ref{ch:chapter_2}. In brief, \cite{Agertz2013} demonstrated that stellar radiation, accounts for a substantial fraction of the energy output of a stellar population. Ionizing photons contribute

\section{Semi-Analytic Models / One-zone Models}

@@

\section{Hydrodynamics}
\textit{(This will be a sub-section of something, but putting here for now as placeholder).}

Astrophysical hydrodynamics simulations are first distinguished by the numerical methods employed to solve the Euler equations that describe the time ($t$) evolution of the energy density ($E$), density ($\rho$), pressure ($p$), and peculiar velocity ($\bm{v}$) of a fluid:
\begin{equation}
  \frac{\delta \rho}{\delta t} + \nabla \cdot \left(\rho \bm{v}\right)  = 0,\\
  \frac{\delta \rho \bm{v}}{\delta t} + \nabla \cdot \left(\rho \bm{v}\bm{v} + \bm{I}p \right) = 0,\\
  \frac{\delta E}{\delta t} + \nabla \cdot \left[\left(E +p\right)\bm{v}\right] = 0.
\end{equation}

Historically, codes fall into one of two camps: 1) Eulerian grid-based codes, including the popular codes / algorithms commonly in use (in some way) today such as \code{Zeus} \citep{StoneNorman1992}, \code{FLASH} \citep{FLASH}, \code{RAMSES} \citep{Teyssier2002}, \code{Athena} \citep{Athena}, \code{ART-II} \citep{Rudd2008}, and \code{Enzo} \citep{Enzo2014}, and 2) particle-based Lagrangian methods known as smooth particle hydrodynamics (SPH), such as \code{Gadget} \citep{Springel2005}, \code{PKDGRAV-2} \citep{Stadel2001}, \code{Gasoline} \citep{Wadsley2004}, and \code{Changa} \citep{Menon2015}. However, recent codes blur the lines between these distinctions with new algorithms for solving the fluid equations on a moving-mesh (e.g. \code{AREPO} \citep{Springel2010}), or meshless finite-mass / finite-volume methods (such as those implemented in \code{Gizmo} \citep{Hopkins2015}). Given the large variance in numerical methods, recent studies have examined the differences between these numerical implementations \citep[e.g.][]{Agertz2007}, including large-scale code comparison projects \citep[e.g.][]{AGORA,AGORA2}. While developing a notion of which code is ``correct" is an ill-posed problem, these studies have allowed for improvement across implementations and an insight into how numerical methods themselves drive uncertainty in our understanding of astrophysical problems.

In this Dissertation, we make use of the adaptive-mesh refinement (AMR) hydrodynamics code \code{Enzo}, as described in greater detail in Chapter~\ref{ch:chapter_1}. As it pertains to galactic chemical evolution, however, the use of a grid-based code has the advantage that metals (which are advected as passive scalars that follow the fluid-flow in the simulations) are allowed to mix and diffuse naturally across fluid elements (grid cells). This is not possible in native SPH implementations, yet is necessary to reproduce realistic galactic chemical evolution properties as shown by multiple recent works testing implementations of diffusion in SPH simulations \citep[e.g.][]{Shen2010,Su2017a,Escala2018}. The disadvantage of the diffusion in \code{Enzo}, however, is that it is entirely numerical, and therefore its exact properties are challenging to characterize and are strongly resolution dependent. Throughout this dissertation we attempt to account for the effects of resolution on our results, but do not explicitly examine the properties of numerical diffusion itself.
